---
title: "Analyzing Stocks Using R"
author: "SW"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This article is adapted from a Medium article [**Analyzing Stocks Using R**](https://nbviewer.org/github/stevenkhwun/Analyzing-Stocks-Using-R/blob/main/Analyzing-Stocks-Using-R_Harjanto.pdf) by _Joy Gracia Harjanto_ publiched at 24 August 2018.

Amazon (AMZN)’s stock experienced a 95.6% (+$918.93) increase this past year,
which makes Amazon (AMZN) a desirable choice for many investors. Many analysts
also believe Amazon (AMZN)’s value will continue to increase in the upcoming years. While it sounds tempting to purchase the stock, an elaborate in-depth analysis should be done to avoid purchasing the stock based on speculation.

I implemented my knowledge in Statistics and R skills to analyze its performance from a technical side and predict its future price. I heavily relied on a packages frequently used in quantitative finance such as `quantmod` and `xts` to fulfill my goal.

Here is the complete list of packages I used:

```{r echo=TRUE, warning=FALSE, message=FALSE}
# Load the required packages
library(quantmod)
library(xts)
library(rvest)
library(tidyverse)
library(stringr)
library(forcats)
library(lubridate)
library(plotly)
library(dplyr)
library(PerformanceAnalytics)
```

I started my analysis by obtaining the log returns of Amazon’s stock beginning August 1st, 2008 to August 17th, 2018:

```{r results='hide'}
getSymbols("AMZN", from="2008-08-01", to="2018-08-17")
AMZN_log_returns <- AMZN %>% Ad() %>% dailyReturn(type='log')
```

The log returns doesn’t make sense at this stage but it is actually the fundamental of my analysis. I will explain later.

## Technical Analysis

I started by doing a technical analysis of Amazon’s stock:

The first chart series graph is straightforward as it shows Amazon’s price chart:

```{r}
AMZN %>% Ad() %>% chartSeries()
```

The second chart series show the Bollinger Band chart, % Bollinger change, Volume
Traded and Moving Average Convergence Diverence for 2018:

```{r}
AMZN %>% chartSeries(TA='addBBands(); addVo(); addMACD()', subset = '2018')
```

The moving average is important to understanding Amazon (AMZN)’s technical charts. It smoothes out daily price fluctuations by averaging stock prices and is effective in identifying potential trends.

The Bollinger Band chart plots two standard deviations away from the moving average and is used to measure the stock’s volatiliy. The Volume chart shows how its stocks are traded on the daily. The Moving Average Convergence Divergence gives technical analysts buy/sell signals. The rule of thumb is: If it falls below the line, it is time to sell. If it rises above the line, it is experiencing an upward momentum.

The charts above are usually used to decide whether to buy/sell a stock. Since I am not a certified financial analyst, I decided to do additional research to convince myself.
