---
title: "Analyzing Stocks Using R"
author: "SW"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This article is adapted from a Medium article [**Analyzing Stocks Using R**](https://nbviewer.org/github/stevenkhwun/Analyzing-Stocks-Using-R/blob/main/Analyzing-Stocks-Using-R_Harjanto.pdf) by _Joy Gracia Harjanto_ publiched at 24 August 2018.

Amazon (AMZN)’s stock experienced a 95.6% (+$918.93) increase this past year,
which makes Amazon (AMZN) a desirable choice for many investors. Many analysts
also believe Amazon (AMZN)’s value will continue to increase in the upcoming years. While it sounds tempting to purchase the stock, an elaborate in-depth analysis should be done to avoid purchasing the stock based on speculation.

I implemented my knowledge in Statistics and R skills to analyze its performance from a technical side and predict its future price. I heavily relied on a packages frequently used in quantitative finance such as `quantmod` and `xts` to fulfill my goal.

Here is the complete list of packages I used:

```{r echo=TRUE, warning=FALSE, message=FALSE}
# Load the required packages
library(quantmod)
library(xts)
library(rvest)
library(tidyverse)
library(stringr)
library(forcats)
library(lubridate)
library(plotly)
library(dplyr)
library(PerformanceAnalytics)
```

I started my analysis by obtaining the log returns of Amazon’s stock beginning August 1st, 2008 to August 17th, 2018:

```{r results='hide'}
getSymbols("AMZN", from="2008-08-01", to="2018-08-17")
AMZN_log_returns <- AMZN %>% Ad() %>% dailyReturn(type='log')
```

```{r}
AMZN %>% Ad() %>% chartSeries()
AMZN %>% chartSeries(TA='addBBands(); addVo(); addMACD()', subset = '2018')
```


You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
