---
title: "Adding Multiple Chart Series in Quantmod R"
author: "SW"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document is based on a post on **stackoverflow** with the title [Adding Multiple Chart Series in Quantmod R](https://stackoverflow.com/questions/38136008/adding-multiple-chart-series-in-quantmod-r).

## Getting the data

We load the required package first:

```{r message=FALSE, warning=FALSE}
library(quantmod)
```

You can get multiple stocks by specifying the required stocks in a vector by `c()` function. Further, we can assign these variables in a created environment, named `data` below.

```{r results=FALSE}
tickers <- c('GLD', 'GDX')
data <- new.env()
getSymbols(tickers, src = 'yahoo', from = '1980-01-01', env = data)
```

However, you need to add `data$` before to read the data, for example:

```{r}
head(data$GDX)
```


## Creating multiple chart series

We use `chart_Series` to create multiple chart series:

```{r eval=FALSE}
chart_Series(Cl(data$GLD))
add_TA(Cl(data$GDX), on = 1)
```

However, in RStudio, the above codes will produce two plots, one with only one series `GLD` and the second one has both series in the same plot. As a result, we need to assign the plot to a varaible first and print it the chart is completed.

```{r}
Chart1 <- chart_Series(Cl(data$GLD))
Chart1 <- add_TA(Cl(data$GDX), on = 1)
Chart1
```

And then if you want RSI below in a sub panel, just add `add_RSI()`.

```{r}
Chart1 <- add_RSI()
Chart1
```

## Compare relative changes in multiple chart series

To see relative changes more easily, you can normalise your price series in many ways. This is a typical approach (using a 0 origin is what Google charts does). First we create the normalising function.

```{r}
# Create the normalising function
normalise_series <- function(xdat) xdat / coredata(xdat)[1]
```

We get one more stock price. In this time, we don't assign it to the `data$` environment. As a result, when we retrieve the data, we just need to type the variable name.

```{r results='hide'}
getSymbols("USO")
```

Further, we create a variable `window` which assign a period to it. In the following example, we assign `2013/` to the window variable, which means that we get the data from 2013 up to now. We can also assign, say `2013/2016-6` to the variable, which means we want the data from 2013 up to June of 2016.

```{r}
window <- "2013/"
```

We can also define a default chart line to chart_Series in `mytheme` object.

```{r}
# Define colour of default chart line to chart_Series in mytheme object
# which is passed to chart_Series
mytheme <- chart_theme()
mytheme$col$line.col <- "darkgreen"
```

Now, we can create the chart:

```{r}
Chart2 <- chart_Series(normalise_series(Cl(data$GLD)[window])-1, theme = mytheme)
Chart2 <- add_TA(normalise_series(Cl(data$GDX)[window]) -1, on = 1, col = "red", lty = 3)
Chart2 <- add_TA(normalise_series(Cl(USO)[window])-1, on = 1, col = "blue", lty = 2)
Chart2
```

Finally, we can add RSI below.

```{r}
Chart2 <- add_TA(RSI(Cl(data$GLD)), on = NA, col = "darkgreen")
Chart2 <- add_TA(RSI(Cl(data$GDX)), on = 2, col = "red", lty = 3)
# Or add RSIs on different subpanels to improve readability of charts:
Chart2 <- add_TA(RSI(Cl(USO)), on = NA, col = "blue", lty = 2)
Chart2
```

