---
title: "Quantitative Finance Applications in R - 3: Plotting xts Time Series"
author: "SW"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document is adapted from an article titled [**Quantitative Finance Applications in R - 3: Plotting xts Time Series**](https://blog.revolutionanalytics.com/2014/01/quantitative-finance-applications-in-r-plotting-xts-time-series.html) by _Daniel Hanson_.


## Introduction and Data Setup

Last time, we included a couple of examples of plotting a single `xts` time series using the `plot(.)` function (ie, said function included in the `xts` package).  Today, we’ll look at some quick and easy methods for plotting overlays of multiple `xts` time series in a single graph.  As this information is not explicitly covered in the examples provided with `xts` and base `R`, this discussion may save you a bit of time.

To start, let’s look at five sets of cumulative returns for the following ETF’s:

* SPY SPDR S&P 500 ETF Trust
* QQQ PowerShares NASDAQ QQQ Trust
* GDX Market Vectors Gold Miners ETF
* DBO PowerShares DB Oil Fund (ETF)
* VWO Vanguard FTSE Emerging Markets ETF

We first obtain the data using quantmod, going back to January 2007:

```{r message=FALSE, results='hide'}
library(quantmod)
tckrs <- c("SPY", "QQQ", "GDX", "DBO", "VWO")
getSymbols(tckrs, from = "2007-01-01")
```

Then, extract just the closing prices from each set:

```{r}
SPY.Close <- SPY[,4]
QQQ.Close <- QQQ[,4]
GDX.Close <- GDX[,4]
DBO.Close <- DBO[,4]
VWO.Close <- VWO[,4]
```

What we want is the set of cumulative returns for each, in the sense of the cumulative value of $1 over time.  To do this, it is simply a case of dividing each daily price in the series by the price on the first day of the series.  As `SPY.Close[1]`, for example, is itself an `xts` object, we need to coerce it to numeric in order to carry out the division:

```{r}
SPY1 <- as.numeric(SPY.Close[1])
QQQ1 <- as.numeric(QQQ.Close[1])
GDX1 <- as.numeric(GDX.Close[1])
DBO1 <- as.numeric(DBO.Close[1])
VWO1 <- as.numeric(VWO.Close[1])
```

Then, it’s a case of dividing each series by the price on the first day, just as one would divide an R vector by a scalar.  For convenience of notation, we’ll just save these results back into the original ETF ticker names and overwrite the original objects:

```{r}
SPY <- SPY.Close/SPY1
QQQ <- QQQ.Close/QQQ1
GDX <- GDX.Close/GDX1
DBO <- DBO.Close/DBO1
VWO <- VWO.Close/VWO1
```

We then merge all of these `xts` time series into a single `xts` object (i.e. a matrix):

```{r}
basket <- cbind(SPY, QQQ, GDX, DBO, VWO)
is.xts(basket)
```

Note that `is.xts(basket)` returns `TRUE`. We can also have a look at the data and its structure:

```{r}
head(basket, n =3)
tail(basket, n =3)
```

Note that we have a few `NA` values here. This will not be of any significant consequence for demonstrating plotting functions, however.

We will now look how we can plot all five series, overlayed on a single graph.  In particular, we will look at the `plot(.)` functions in both the `zoo` and `xts` packages.


## Using `plot(.)` in the zoo package

The `xts` package is an extension of the `zoo` package, so coercing our `xts` object basket to a zoo object is a simple task:

```{r}
zoo.basket <- as.zoo(basket)
```

Looking at `head(zoo.basket)` and `tail(zoo.basket)`, we will get output that looks the same as what we got for the original `xts` `basket` object, as shown above; the date to data mapping is preserved. The `plot(.)` function provided in zoo is very simple to use, as we can use the whole `zoo.basket` object as input, and the `plot(.)` function will overlay the time series and scale the vertical axis for us with the help of a single parameter setting, namely the screens parameter.

Let’s now look at the code and the resulting plot in the following example, and then explain what’s going on:

```{r}
# Set a color scheme:
tsRainbow <- rainbow(ncol(zoo.basket))

# Plot the overlayed series
plot(x = zoo.basket, ylab = "Cumulative Return", main = "Cumulative Returns",
     col = tsRainbow, screens = 1)

# Set a legend in the upper left hand corner to match color to return series
legend(x = "topleft", legend = c("SPY", "QQQ", "GDX", "DBO", "VWO"),
       lty = 1,col = tsRainbow)
```

We started by setting a color scheme, using the `rainbow(.)` command that is included in the base R installation.  It is convenient as R will take in an arbitrary positive integer value and select a sequence of distinct colors up to the number specified.  This is a nice feature for the impatient or lazy among us (yes, guilty as charged) who don’t want to be bothered with picking out colors and just want to see the result right away.

Next, in the `plot(.)` command, we assign to `x` our “matrix” of time series in the `zoo.basket` object, labels for the horizontal and vertical axes (`xlab`, `ylab`), a title for the graph (`main`), the the colors (`col`). Last, but crucial, is the parameter setting `screens = 1`, which tells the plot command to overlay each series in a single graph.

Finally, we include the `legend(.)` command to place a color legend at the upper left hand corner of the graph.  The position (`x`) may be chosen from the list of keywords "`bottomright`", "`bottom`", "`bottomleft`", "`left`", "`topleft`", "`top`", "`topright`", "`right`" and "`center`"; in our case, we chose "`topleft`".  The `legend` parameter is simply the list of ticker names.  The `lty` parameter refers to “line type”, and by setting it to `1`, the lines in the legend are shown as solid lines, and as in the `plot(.)` function, the same color scheme is assigned to the parameter `col`.

Back to the color scheme, we may at some point need to show our results to a manager or a client, so in that case, we probably will want to choose colors that are easier on the eye.  In this case, one can just store the colors into a vector, and then use it as an input parameter.  For example, set

```{r}
myColors <- c("red", "darkgreen", "goldenrod", "darkblue", "darkviolet")
```

Then, just replace `col = tsRainbow` with `col = myColors` in the `plot` and `legend` commands:

```{r}
plot(x = zoo.basket, xlab = "Time", ylab = "Cumulative Return",
     main = "Cumulative Returns", col = myColors, screens = 1)
legend(x = "topleft", legend = c("SPY", "QQQ", "GDX", "DBO", "VWO"),
       lty = 1, col = myColors)
```

## Using `plot(.)` in the `xts` package

While the `plot(.)` function in `zoo` gave us a quick and convenient way of plotting multiple time series, it didn’t give us much control over the scale used along the horizontal axis. Using `plot(.)` in `xts` remedies this; however, it involves doing more work.  In particular, we can no longer input the entire “matrix” object; we must add each series separately in order to layer the plots. We also need to specify the scale along the vertical axis, as in the `xts` case, the function will not do this on the fly as it did for us in the zoo case.

We will use individual columns from our original `xts` object, `basket`.  By using `basket` rather than `basket.zoo`, this tells R to use the `xts` version of the function rather than the `zoo` version (i.e. an overloaded function in traditional object oriented programming).  Let’s again look at an example and the resulting plot, and then discuss how it works:


```{r fig.keep='last'}
plot(x = basket[,"SPY.Close"], xlab = "Time", ylab = "Cumulative Return",
     main = "Cumulative Returns", ylim = c(0.0, 12), major.ticks= "years",
     minor.ticks = FALSE, col = "red")
lines(x = basket[,"QQQ.Close"], col = "darkgreen")
lines(x = basket[,"GDX.Close"], col = "goldenrod")
lines(x = basket[,"DBO.Close"], col = "darkblue")
lines(x = basket[,"VWO.Close"], col = "darkviolet")
addLegend('topleft', legend.names = c("SPY", "QQQ", "GDX", "DBO", "VWO"),
          lty = 1, col = myColors)
```

As mentioned, we need to add each time series separately in this case in order to get the desired overlays.  If one were to try `x` = basket in the plot function, the graph would only display the first series (`SPY`), and a warning message would be returned to the R session.  So, we first use the `SPY` series as input to the `plot(.)` function, and then add the remaining series with the `lines(.)` command.  The color for each series is also included at each step (the same colors in our `myColors` vector).

As for the remaining arguments in the `plot` command, we use the same axis and title settings in `xlab`, `ylab`, and `main`.  We set the scale of the vertical axis with the `ylim` parameter; noting from our previous example that `VWO` hovered near zero at the low end, and that `DBO` reached almost as high as 12, we set this range from 0.0 to 12.  Two new arguments here are the `major.ticks` and `minor.ticks` settings. The `major.ticks` argument represents the periods in which we wish to chop up the horizontal axis; it is chosen from the set

`{"years", "months", "weeks", "days", "hours", "minutes", "seconds"}`

In the example above, we chose "`years`".  The `minor.ticks` parameter can take values of `TRUE`/`FALSE`, and as we don’t need this for the graph, we choose `FALSE`.  The same `legend` command that we used in the `zoo` case can be used here as well (using `myColors` to indicate the color of each time series plot).

## Wrap-up

A new package, called `xtsExtra`, includes a new `plot(.)` function that provides added functionality, including a legend generator.  However, while it is available on R-Forge, it has not yet made it into the official CRAN repository.  More sophisticated time series plotting capability can also be found in the `quantmod` and `ggplot2` packages, and we will look at the `ggplot2` case in an upcoming post.  However, for plotting `xts` objects quickly and with minimal fuss, the `plot(.)` function in the `zoo` package fills the bill, and with a little more effort, we can refine the scale along the horizontal axis using the `xts` version of `plot(.)`.  `R` help files for each of these can be found by selecting `plot.zoo` and `plot.xts` respectively in help searches.

## Further Notes

The following notes on using `plot.xts` is adapted from a [article](https://stackoverflow.com/questions/69498375/in-r-how-can-i-add-a-legend-to-a-plot-of-an-xts-object) from **stackoverflow**.

```{r message=FALSE, warning=FALSE}
# Import data
quantmod::getSymbols(c('SPY','IWM'),
                     src  = 'yahoo', 
                     from = '2010-01-01', 
                     to   = '2021-02-01',
                     auto.assign = TRUE, 
                     warnings = FALSE)
close <- cbind(SPY$SPY.Close, IWM$IWM.Close)
```

Plot the data:

```{r fig.keep='last'}
plot.xts(SPY$SPY.Close, ylim = c(0,500), main = "", col = "red")
lines(IWM$IWM.Close, col = "blue")
addLegend("topleft", on=1, 
          legend.names = c("SPY", "IWM"), 
          lty=c(1, 1), lwd=c(2, 1),
          col=c("red", "blue"))
```


