---
title: "Retrieve economic data using `quantmod`"
author: "SW"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Various financial and economic data can be retrieved by using `quantmod` library. Of course, we need to load the library first:

```{r message=FALSE}
# Load the 'quantmod' library
library(quantmod)
```


## Retrieve stock prices from Yahoo Finance

### For stocks with stock tickers beginning with letter

We demonstrate the method by obtaining the stock prices of Apple Inc. (AAPL).

```{r results='hide'}
# Get stock prices of Apple
getSymbols("AAPL", from=(Sys.Date()-730), to=Sys.Date())
```

We can observe the loaded data with the following commands:

```{r}
# The size of the loaded dataset
dim(AAPL)
# The first 6 rows of the dataset
head(AAPL)
# The last 6 rows of the dataset
tail(AAPL)
```

We can create a plot of closing prices and trading volumes with the following commands:

```{r}
# Create time plot
# The option 'theme="white" set the background of the time plot
# The default of this option is "black"
chartSeries(AAPL, theme = "white")
```

### For stocks with stock tickers beginning with number

We need to set the option `auto.assign=FALS` it the stock tickers beginning with a number when calling `getSymbols()`. This will allow you to assign the data returned to a variable name of your choice.

We demonstrate the method by obtaining the stock prices of HSBC (0005).

```{r results='hide', warning=FALSE}
# Get HSBC stock price
HSBC.HK <- getSymbols("0005.HK",from="2022-08-17", to=Sys.Date(),
                      auto.assign = FALSE)
```

We can observe the loaded data with the following commands:

```{r}
# The size of the loaded data set
dim(HSBC.HK)
# The last 3 rows of data
tail(HSBC.HK, n = 3)
```

We can create a plot of closing prices and trading volumes with the following commands:

```{r}
# Create time plot
chartSeries(HSBC.HK,theme="white")
```

We now try another example in getting stock prices of Hon Hai Precision Industry Co., Ltd. (2317.TW) listed in Taiwan Stock Exchange.

```{r warning=FALSE}
# Get Hon Hai Precision Industry Co., Ltd
Honhai.TW <- getSymbols("2317.TW",from=(Sys.Date()-730), to=Sys.Date(),
                      auto.assign = FALSE)
tail(Honhai.TW, n = 3)
```

We can compare performance of the stocks by the following commands:

```{r}
normalise_series <- function(xdat) xdat / coredata(xdat)[1]
window <- "2023/"
mytheme <- chart_theme()
mytheme$col$line.col <- "darkgreen"
TA <- chart_Series(normalise_series(Cl(AAPL)[window])-1, name = "AAPL vs HSBC", legend = "AAPL", theme = mytheme)
TA <- add_TA(normalise_series(Cl(HSBC.HK)[window])-1, on = 1, legend = "HSBC", col = "red", lty = 3)
TA
```

## Retrieve financial information using `getQuote`

Please see the following example:

```{r}
what_metrics <- yahooQF(c("Name",
                          "Volume",
                          "P/E Ratio",
                          "Dividend Yield"))
tickers <- c("AAPL", "0005.HK")
getQuote(tickers, what = what_metrics)
```

You can get the list of available financial information by typing:

```{r eval=FALSE}
yahooQF()
```



## Retrieve daily interest rate of 10-year treasures notes

```{r}
# Retrieve daily interest rate of 10-year treasures notes
# These are interest rates from the Chicago Board Options Exchange (CBOE)
# obtained from Yahoo Finance
getSymbols("^TNX")
#See the size of the loaded data set
dim(TNX)
#See the first 6 row of data
#Note that the trading volume is 0
head(TNX)
#See the last 6 row of data
tail(TNX)
#Obtain time plot of daily interest rates without trading volume
#The subcommand TA=NULL is used to omit the time plot of volume
chartSeries(TNX,theme="white",TA=NULL)
```


## Retrieve economic data from the Federal Reserve Economic Data (FRED) of Federal Reserve Bank of St. Louis

```{r}
# Retrieve financial data from FRED
# Load monthly unemployment rate from FRED
getSymbols("UNRATE", src="FRED")
#See the size of the loaded data set
dim(UNRATE)
#See the first 6 row of data
head(UNRATE)
#See the last 6 row of data
tail(UNRATE)
#Obtain time plot of monthly U.S. unemployment rates
#The subcommand theme="white" set the background of the time plot
#The default is black
chartSeries(UNRATE)
```

