---
title: "Analyzing Stocks Using R"
author: "SW"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale("LC_ALL", "English")
```


```{r echo=TRUE, warning=FALSE, message=FALSE}
# Load the required packages
library(quantmod)
library(xts)
library(rvest)
library(tidyverse)
library(stringr)
library(forcats)
library(lubridate)
library(plotly)
library(dplyr)
library(PerformanceAnalytics)
```


## Investment rules

### IBD Methodology

* Invest in stocks with recent quarterly and annual earnings growth of at least 25%.
* Look for companies that have new, game-changing products and services.
* Consider not-yet-profitable companies, often IPOs, that are generating tremendous revenue growth.


### Keep track of the market

* Market is in a confirmed uptrend and move to cash when stock market goes into a correction.


### Sell signals

* Any stocks that falls 7% or 8% from your purchase price should be jettisoned.
* beware of sharp breaks below the 50-day or 10-week moving average



## Watching List

Firstly, I create my watching list:

### AI related

```{r warning=FALSE}
# Get raw data
# Type "yahooQF()" to get a list of metrics
ai_tickers <- c("ORCL", "CRM", "PLTR", "V", "UNH", "JPM", "MA", "TM", "SIEGY",
                "LLY", "NVO", "JNJ", "NVDA", "AMD", "ARM", "TSM", "TSLA",
                "UBER", "CKI.L")

what_metrics1 <- yahooQF(c("Name",
                          "Currency",
                          "Last Trade (Price Only)",
                          "52-week Range",
                          "50-day Moving Average",
                          "P/E Ratio",
                          "Dividend Yield"))
ai_stocks <- getQuote(ai_tickers, what = what_metrics1)

# Convert to a more readable format
ai_stocks[, "Yield (%)"] <- ai_stocks[, "Dividend Yield"] * 100
ai_stocks[, "Yield (%)"] <- sprintf(ai_stocks[, "Yield (%)"], fmt = '%.2f')
ai_stocks[, "P/E Ratio"] <- sprintf(ai_stocks[, "P/E Ratio"], fmt = '%.2f')
ai_stocks[, "Trade Time"] <- format(ai_stocks[, "Trade Time"], format = "%b %d, %Y")

# Print the table
library(knitr)
kable(ai_stocks[, c(1:7, 9)], caption = "Latest Performance")
```


### Others

```{r warning=FALSE}
# Get raw data
# Type "yahooQF()" to get a list of metrics
ot_tickers <- c("WMT", "COST", "SNA", "SWK", "6586.T")

what_metrics1 <- yahooQF(c("Name",
                          "Currency",
                          "Last Trade (Price Only)",
                          "52-week Range",
                          "50-day Moving Average",
                          "P/E Ratio",
                          "Dividend Yield"))
ot_stocks <- getQuote(ot_tickers, what = what_metrics1)

# Convert to a more readable format
ot_stocks[, "Yield (%)"] <- ot_stocks[, "Dividend Yield"] * 100
ot_stocks[, "Yield (%)"] <- sprintf(ot_stocks[, "Yield (%)"], fmt = '%.2f')
ot_stocks[, "P/E Ratio"] <- sprintf(ot_stocks[, "P/E Ratio"], fmt = '%.2f')
ot_stocks[, "Trade Time"] <- format(ot_stocks[, "Trade Time"], format = "%b %d, %Y")

# Print the table
library(knitr)
kable(ot_stocks[, c(1:7, 9)], caption = "Latest Performance")
```








```{r results='hide'}
index <- c("^GSPC", "^IXIC", "^DJI", "^FTSE", "^N225","^TWII", "^HSI")

us_tickers <- c("AAPL", "AMZN", "NVDA", "GOLD", "GOOG", "GOOGL", "BLK")

hk_tickers <- c("0001.HK", "0005.HK", "0939.HK")
```

Get some knowledge of their performance first:

```{r warning=FALSE}
# Get raw data
# Type "yahooQF()" to get a list of metrics
what_metrics1 <- yahooQF(c("Name",
                          "Currency",
                          "Last Trade (Price Only)",
                          "52-week Range",
                          "50-day Moving Average",
                          "P/E Ratio",
                          "Dividend Yield"))
stocks <- getQuote(us_tickers, what = what_metrics1)

# Convert to a more readable format
stocks[, "Yield (%)"] <- stocks[, "Dividend Yield"] * 100
stocks[, "Yield (%)"] <- sprintf(stocks[, "Yield (%)"], fmt = '%.2f')
stocks[, "P/E Ratio"] <- sprintf(stocks[, "P/E Ratio"], fmt = '%.2f')
stocks[, "Trade Time"] <- format(stocks[, "Trade Time"], format = "%b %d, %Y")

# Print the table
library(knitr)
kable(stocks[, c(1:7, 9)], caption = "Latest Performance")
```


```{r warning=FALSE}
# Get raw data
# Type "yahooQF()" to get a list of metrics
hkstocks <- getQuote(hk_tickers, what = what_metrics1)

# Convert to a more readable format
hkstocks[, "Yield (%)"] <- hkstocks[, "Dividend Yield"] * 100
hkstocks[, "Yield (%)"] <- sprintf(hkstocks[, "Yield (%)"], fmt = '%.2f')
hkstocks[, "P/E Ratio"] <- sprintf(hkstocks[, "P/E Ratio"], fmt = '%.2f')
hkstocks[, "Trade Time"] <- format(hkstocks[, "Trade Time"], format = "%b %d, %Y")

# Print the table
library(knitr)
kable(hkstocks[, c(1:7, 9)], caption = "Latest Performance")
```



```{r}
what_metrics2 <- yahooQF(c("Name",
                          "Last Trade (Price Only)",
                          "52-week Range",
                          "Currency",
                          "Volume"))
```

```{r}
getSymbols("^HSI", from="1997-07-02")
head(HSI, n=5)
tail(HSI, n=3)
```

```{r}
window <- "2021/"
chartSeries((HSI)[window],theme="white")
```

```{r}
HSI %>% chartSeries(TA='addBBands(); addVo(); addMACD()', subset = '2023/', theme = "white")
```

```{r}
HSBC.HK <- getSymbols("0005.HK",from="1997-08-17", to=Sys.Date(),
                      auto.assign = FALSE)
HSBC.HK %>% chartSeries(TA='addBBands(); addSMA(n=250); addVo(); addMACD()', subset = '2023/', theme = "white")
```





```{r}
window <- "2002/"
chart_Series(Vo(HSI)[window])
```


```{r}
getSymbols("AMZN")
getSymbols(us_tickers)
tail(AMZN, n=3)
```


The log returns doesn’t make sense at this stage but it is actually the fundamental of my analysis. I will explain later.

## Technical Analysis

I started by doing a technical analysis of Amazon’s stock:

The first chart series graph is straightforward as it shows Amazon’s price chart:

```{r}
GOLD %>% chartSeries(TA='addBBands(); addVo(); addMACD()', subset = '2023/', theme = "white")
```

```{r}
NVDA %>% chartSeries(TA='addBBands(); addSMA(n=250); addVo(); addMACD()', subset = '2023/', theme = "white")
```


The second chart series show the Bollinger Band chart, % Bollinger change, Volume
Traded and Moving Average Convergence Diverence for 2018:

```{r}
HSI %>% chartSeries(TA='addBBands(); addVo(); addMACD()', subset = '2023/', theme = "white")
```



The moving average is important to understanding Amazon (AMZN)’s technical charts. It smoothes out daily price fluctuations by averaging stock prices and is effective in identifying potential trends.

The Bollinger Band chart plots two standard deviations away from the moving average and is used to measure the stock’s volatility. The Volume chart shows how its stocks are traded on the daily. The Moving Average Convergence Divergence gives technical analysts buy/sell signals. The rule of thumb is: If it falls below the line, it is time to sell. If it rises above the line, it is experiencing an upward momentum.

The charts above are usually used to decide whether to buy/sell a stock. Since I am not a certified financial analyst, I decided to do additional research to convince myself.
